---
title: "REddyProc for Eddy Covariance Data Analysis"
author: "Yusri Yusup"
date: "`r Sys.Date()`"
output: 
  powerpoint_presentation:
    reference_doc: ../template/my_template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Acknowledgements

This content was built on top of the talk by Thomas Wutzler. Credit goes to him for his: 

1. Presentation at (https://www.youtube.com/watch?v=-b0vc4u8kls&t=354s) and 
2. Material at https://github.com/bgctw/EGU19EddyCourse/blob/master/Source/DEGebExample_complete.Rmd.

You can view more details about the package at https://bgc.iwww.mpg.de/5622399/REddyProc.

# Prerequisites

## Knowledge

It would be easier for you to follow the talk if you have brief backgrounds on:

1. R language
2. Eddy covariance
3. Net Ecosystem Exchange

## Software

You would need the software below to perform the steps discussed in the talk.

1. R 
2. RStudio
3. Packages: REddyProc

# Introduction to the REddyProc Package: When Do We Use It?

The REddyProc package become useful after the pre-processing of the raw, high-frequency eddy covariance data. 

Using software, such as EddyPro, the data was already quality-checked using protocols, e.g., 0-1-2, 0-9 quality flag systems, etc.

We use this package to:

1. Post-process the Net Ecosystem Exchange (NEE) data.
2. Estimate NEE, e.g., the annual, seasonal sums.
3. Determine the contributions of processes within NEE, e.g., production and respiration.


# Steps in Using REddyProc

There are three main steps in a typical REddyProc analysis workflow. 

1. Check for under-developed turbulence or low friction velocity ($u_*$).

    * The package can suggest threshold values of $u_*$.

2. Fill gaps introduced due to the discarding of fluxes below the thresholds.

3. Separating fluxes into groups: Gross Primary Production (GPP) and Ecosystem Resipiration ($R_{eco}$).


# Learning Outcomes

At the end of the walk-through, you would be able to:

1. Understand the functions of the REddyProc package.
2. Explain the analysis steps of the package.

# Step 0: Preliminary Work

## Install the Package

The package needs to be installed prior to use. You might also need to install other packages  to run REddyProc. You can do so by using the `install.packages` command for their installation.

```{r Install Package, echo=TRUE}
install.packages("REddyProc", repos = "http://cran.us.r-project.org")
```

You will need to load the package after a successful installation. 

```{r Load Package, echo=TRUE}
library(REddyProc)
```


# Step 1: Prepare the Data

## Step 1-1: Import the Data

The eddy covariance data that will be used in this walk-through are included in the package. They are for demo purposes.

1. `DEGebExample`, the Gebesee, Germany, data from 2004 to 2006. 
2. `Example_DETha98`, the Tharandt, Germany, data for the year 1998. 

The full data is downloadable at http://www.europe-fluxdata.eu/home/ after registration.

Load the Gebesee data.

```{r Load Gebesee Data, echo = TRUE}
data(DEGebExample)
```

Load the Tharandt data.

```{r Load Tharandt Data, echo=TRUE}
data(Example_DETha98)
```

# Data Overview: Gebesee

Get an overview of the data. Look at the data parameters and take note of missing data or `NA`.

## The Gebesee, Germany, Data 

### Characteristics

* Surface: Agriculture
* Time zone: +1 GMT
* Latitude, Longitude: 51.1N, 10.9E


---

Notice that VPD is not in the dataset.

```{r Summary Gebesee, echo=TRUE}
summary(DEGebExample)
```


# Data Overview: Tharandt

## The Tharandt, Germany, Data

### Characteristics

* Surface: Forest
* Time zone: +2 GMT
* Latitude, Longitude: 51.0N, 13.6E

---

Note that the timestamp is not in REddyProc-usable format.

```{r Summary Tharandt, echo=TRUE} 
names(Example_DETha98)
head(Example_DETha98)
```



# Important Parameters

Parameters required for REddyProc are:

1. `DateTime` in the POSIX format.
2. `NEE` or carbon dioxide flux.
3. `Ustar` or friction velocity
4. Meteorological data for the gap-filling and partitioning steps.
      * `Rg`, 
      *`Tair`, 
      * `rH`, and or 
      * `VPD` 

Note that:

* `Rg` is global solar radiation.
* `Tair` is air temperature.
* `rH` is relative humidity.
* `VPD` is vapor pressure deficit.


# Step 1-2: Calculate Needed Parameters

Essential parameters can be calculated from existing parameters using functions available in REddyProc.

Some useful functions are:

1. `fConvertTimeToPosix`. We will use this in the demo.
2. `fCalcVPDfromRHandTair`. We will use this in the demo.
3. `fCalcETfromLE` 
4. `fConvertCtoK`

There are other functions in the package and the function name begins with the prefix `f`.


# Step 1-3-1: Tharandt Dataset: Addressing the Unsupported Timestamp Format

In the Tharandt dataset, the date-time columns are not suitable for REddyProc. It needs to be converted to the `POSIX` format.

```{r Tharandt Timestamp, echo=TRUE}
head(Example_DETha98)
```

# Convert the Timestamp to `POSIX`

We can convert the timestamp using the `fConvertTimeToPosix` function. It will add the `DateTime` column into the data frame.

The `'YDH'` means Year-Day-Hour, and the `Year`, `Day`, and `Hour` arguments require the columns that contains the Year, DoY, and Hour information.

You can find other valid time configurations in the documentation by running the command `?fConvertTimeToPosix`.

---

```{r Tharandt Timestamp Conversion, echo=TRUE}
Example_DETha98V1 <- fConvertTimeToPosix(Example_DETha98, TFormat = c('YDH'),
                    Year = 'Year',
                    Day = 'DoY', 
                    Hour = 'Hour')
head(Example_DETha98V1)
```


# Step 1-3-2: Missing VPD in the Gebesee Data

The Gebesee dataset does not have the VPD parameter, which could be useful for gap-filling and partitioning.

```{r Gebesee Missing VPD, echo=TRUE}
head(DEGebExample)
```


# Calculate VPD

We can calculate VPD using the function `fCalcVPDfromRHandTair`. The input arguments' units are stated in the documentation, `?fCalcVPDfromRHandTair`.

```{r Gebesee Add VPD, echo=TRUE}
VPD <- fCalcVPDfromRHandTair(DEGebExample$rH,   # The unit is %
                             DEGebExample$Tair) # The unit is degree Celsius
DEGebExampleV1 <- cbind(DEGebExample,VPD)
rm(VPD) # A house-keeping step.
head(DEGebExampleV1)
```

# Step 2: Create the Gebesee REddyProc Object Class

Before REddyProc can work on your data, the data has to be converted to the REddyProc object.

Create the data object for the Gebesee data. The ID is `DE-Geb` and the parameters are:

1. `NEE` 
2. `Rg`
3. `Tair`
4. `VPD`
5. `Ustar`

```{r Gebesee Create Class, echo=TRUE}
EProcDEGeb <- sEddyProc$new('DE-Geb', DEGebExampleV1, c('NEE','Rg','Tair','VPD','Ustar'))
```

# Check the Object

Check the additional info of the data.

```{r Gebesee Info Check, echo=TRUE}
EProcDEGeb$sLOCATION
```

---

Add the location information. This is important for the daytime-nighttime partitioning analysis because it requires the time to be accurate.

```{r Gebesee Add Info, echo=TRUE}
EProcDEGeb$sSetLocationInfo(LatDeg = 51.1, LongDeg = 10.9, TimeZoneHour = 1)  
EProcDEGeb$sLOCATION
```


# Step 2: Create the Tharandt REddyProc Object Class
 
Create the class for the Tharandt data. The ID is `DE-Tha` and the parameters are:

1. Net Ecosystem Exchange: `NEE` 
2. Global Radiation: `Rg`
3. Air temperature: `Tair`
4. Vapor pressure deficit: `VPD`
5. Friction velocity: `Ustar`

```{r Tharandt Create Class, echo=TRUE}
EProcDETha <- sEddyProc$new('DE-Tha', Example_DETha98V1, c('NEE','Rg','Tair','VPD','Ustar'))
```

# Check the Object

Check the additional info of the data.

```{r Tharandt Info Check, echo=TRUE}
EProcDETha$sLOCATION
```

---

Add the location information.

```{r Tharandt Add Info, echo=TRUE}
EProcDETha$sSetLocationInfo(LatDeg = 51.0, LongDeg = 13.6, TimeZoneHour = 2)
EProcDETha$sLOCATION
```



# Step 3: u*-Threshold Estimation

Friction velocity, or $u_*$, varies seasonally at Gebesee. Thus, the $u_*$-threshold needs to be estimated for each season. We do this because $u_*$ changes with surface cover. 

A previous study determined the days on which the seasons and $u_*$ shifted. It can be determined by the visual inspection of the data.

| Year | Start Day    |
| :--- | :----------: |
| 2004 | 70, 210, 320 |
| 2005 | 70, 180, 320 |
| 2006 | 120, 350     |


# Step 3-1-2: Adding the Start Days for the Gebesee Data.

Create a data frame for the start days.

```{r Gebesee Adding Start Days, echo=TRUE}
df_startDays <- data.frame(day=c(70,210,320,70,180,320,120,305),
                           year=c(2004,2004,2004,2005,2005,2005,2006,2006))
df_startDays
```

# Creating the Seasonal Factors for Row-Tagging

We can use `usCreateSeasonFactorYdayYear` to change the `df_startDays` data frame to a factor vector that contains values that tag each rows to their respective seasons.

Create the factor vector. 

Note that the product `15*60` is used to make the time be between 00:00 and 00:30. 

The `summary` shows that there are 3312 observations for season `2004001`, i.e., between days 1 and 70, etc.

```{r Gebesee Season Factor, echo=TRUE}
seasonFactor <- usCreateSeasonFactorYdayYear(DEGebExampleV1$DateTime - 15*60,
                                             starts = df_startDays)
```

---

```{r}
summary(seasonFactor)
head(seasonFactor)
```


# Optional: Viewing the Gebesee Data with Season Demarcation-Lines

Because the start days data frame is a collection of integers, we need to change it to the POSIX format.

Create timestamps in the POSIX format from `df_startDays`. Here, we embed a new data frame with the additional column `Hour` into the `fConvertTimeToPosix` function call. The hour is set at `0.25` to be between 00:00 and 00:30, i.e., 00:15.

---

```{r Gebesee Optional Convert df_startDays to POSIX, echo=TRUE}
seasonStartsDate <- fConvertTimeToPosix(data.frame(Year = df_startDays$year, 
                                                   DoY = df_startDays$day, 
                                                   Hour = 0.25), 
                                        TFormat = 'YDH', 
                                        Year = "Year", 
                                        Day = "DoY", 
                                        Hour = "Hour")
seasonStartsDate

```

---

You can check the dates by plotting them on the time series.

```{r Gebesee Optional Plot Season Starting Dates, echo=TRUE}
plot(DEGebExample$DateTime, DEGebExample$NEE, pch=19, xlab = "Time", ylab = 'NEE', 
     cex = 0.1, col = "darkblue")
abline(v = seasonStartsDate$DateTime)
```


# Step 3-2-2: Calculate the u*-Thresholds Distributions

We will estimate the ($u_*$) limits using the `sEstimateUstarScenarios` function. The function will write to the data object. The `seasonFactor` is needed here to tell `REddyProc` the season intervals that it must estimate the $u_*$ thresholds. 

The $u_*$ threshold estimation uses the `usEstUstarThreshold` function, which requires the `NEE`, `Tair`, and `seasonFactor`. The function returns the median value.

# The u*-Threshold Distributions

In this example, the $u_*$-threshold is estimated, using the `usEstUstarThreshold` function, 30 times, and the $u_*$ limits are reported using the default quantiles of 5%, 50%, and 95%: The low, median, and high values of $u_*$-threholds.

The function adds data to the REddyProc object. It creates the $u_*$ scenarios and place it in the object.

```{r Gebesee Calculate u*-Thresholds, echo=TRUE}
EProcDEGeb$sEstimateUstarScenarios(seasonFactor = seasonFactor, 
                                   nSample = 30,
                                   probs = c(0.05,0.50,0.95))
```

# Viewing the Results

The function `sGetEstimatedUstarThresholdDistribution` displays the results.

Useful functions for handling the data in the object:

1. `sExportData`: Export class internal sDATA data frame.
2. `sExportResults`: Export class internal sTEMP data frame with result columns. We can use this after gap-filling the data.

Note that you can create the plots of NEE versus ($u_*$) by using the function `sPlotNEEVersusUStarForSeason`.


---

```{r Gebesee View u*-Thresholds, echo=TRUE}
EProcDEGeb$sGetEstimatedUstarThresholdDistribution()
```


# Step 4-1: Gap-Filling the Gebesee Data

## Step 4-1-1: Check the Use of Seasonal ($u_*$) Thresholds

First, we have to ensure the use of seasonal $u_*$-thresholds. If it is not set in the previous step, check that it is used now.

---

Show the default thresholds: annual

```{r Gebesee Show Default Thresholds, echo=TRUE}
EProcDEGeb$sGetUstarScenarios()
```

---

Instruct REddyProc to use the seasonal thresholds.

```{r Gebesee Use Seasonal Thresholds, echo=TRUE}
EProcDEGeb$useSeaonsalUStarThresholds()
```

---

Confirm that the seasonal thresholds are used by displaying it.

```{r Gebesee Confirm Thresholds, echo=TRUE}
EProcDEGeb$sGetUstarScenarios()
```


# Step 4-1-2: Gap-Fill the Gebesee Data

Gap-fill the data using the function `sMDSGapFillUStarScens`. It will filter the data using the $u_*$-thresholds and gap-fill it. 

`MDS` means Marginal Distribution Sampling, which combines:

1. the Look Up Table (LUT)
2. Mean Diurnal Course (MDC)

Quality flags are created for the gap-filled data: 

* 0: original data
* 1: good quality gap-filled data, i.e., *more parameters* and *shorter time-windows* used.
* More than 1: low quality, i.e., *less parameters* and *longer time-windows* used.

The function also calculates the random error for non-gap records by replacing the original values with gap-filled values.

```{r Gebesee Gap-Filling, message = FALSE, echo=TRUE}
EProcDEGeb$sMDSGapFillUStarScens("NEE", FillAll = TRUE)
```

# Check the New Columns

Check the columns created. Examples are:

* `NEE_05_f`
* `NEE_95_fall`
* `NEE_50_fqc`

Definitions:

* NEE_<scenario>_f: gaps replaced by modeled values (gap-filled).
* NEE_<scenario>_fall: all NEE replaced by modeled values.
* NEE_<scenario>_fqc: quality flag: 0 observations, 1 good quality of gap-filling. 
* The non-bootstrapped data has the `uStar` suffix. 
* The bootstrapped data has the scenario suffix, e.g., `U50`, `U95`, etc.

---

```{r Gebesee Column Names, echo=TRUE}
colnames(EProcDEGeb$sExportResults())
```

# View Some Columns

Plotting a column of the REddyProc object.

```{r Gebesee Summary After Gap-Filling, echo=TRUE}

plot(EProcDEGeb$sDATA$sDateTime, EProcDEGeb$sExportResults()$Ustar_uStar_Thres, pch = 19,
     xlab = 'Time', ylab = 'u*-threshold')


plot(DEGebExampleV1$DateTime,DEGebExampleV1$NEE, pch = 19, 
     cex = 0.1, col = "darkblue",
     xlab = 'Time', ylab = 'NEE', ylim=c(-50,20), 
     main = "Before u*-Filtering")
plot(EProcDEGeb$sDATA$sDateTime, EProcDEGeb$sExportResults()$NEE_uStar_orig, 
     pch = 19, cex = 0.1, col = "darkblue",
     xlab = 'Time', ylab = 'NEE_uStar_orig', ylim=c(-50,20), 
     main = "NEE After u*-Filtering")
plot(EProcDEGeb$sDATA$sDateTime, EProcDEGeb$sExportResults()$NEE_U50_f, pch = 19, 
     cex = 0.1, col = "darkblue",
     xlab = 'Time', ylab = 'NEE_U50_f', ylim=c(-50,20), 
     main = "NEE After Gap-Filling")
plot(EProcDEGeb$sDATA$sDateTime, EProcDEGeb$sExportResults()$NEE_U50_fall, pch = 19, 
     cex = 0.1, col = "darkblue",
     xlab = 'Time', ylab = 'NEE_U50_fall', ylim=c(-50,20), 
     main = "NEE After Gap-Filling All")
```


# Step 4-1-3: Gebesee Fingerprint Plot

We can also generate a fingerprint plot using the function `sPlotFingerprintY`. This is for the `NEE_U50_f` parameter and the year 2004.

```{r Gebesee Fingerplot for 2006, echo=TRUE}
EProcDEGeb$sPlotFingerprintY('NEE_uStar_orig', Year = 2006)
EProcDEGeb$sPlotFingerprintY('NEE_U50_f', Year = 2006)
EProcDEGeb$sPlotFingerprintY('NEE_U50_fall', Year = 2006)
```

We can also produce PDF files with legend for all years in sub-directory "figs."

```{r Gebesee Fingerplot for All, echo=TRUE}
EProcDEGeb$sPlotFingerprint('NEE_U50_f', Dir = "../figs")
```

# Step 5-1: Gebesee Preparing the Data for Partitioning

This step requires the data to have the location (lat, lon) and time zone info because REddyProc uses time to estimate day and night hours. We already did this in the *Step 2*.

There are some weather values that are missing and can be gap-filled here. However, we do not need to replace the original values with gap-filled values because we are not going to calculate random error, `FillAll = FALSE`.

```{r Gebesee Gap-Fill Met Data, message=FALSE, echo=TRUE}
EProcDEGeb$sMDSGapFill('Rg', FillAll = FALSE)     
EProcDEGeb$sMDSGapFill('Tair', FillAll = FALSE)     
EProcDEGeb$sMDSGapFill('VPD', FillAll = FALSE)     
```

---

## Step 5-1-1: Gebesee Reichstein Partitioning

In this part, we will partition the data into fractions of the Gross Primary Production (GPP) and ecosystem respiration ($R_{eco}$) using all $u_*$ scenarios. This uses the 'sMRFluxPartitionUStarScens` function. 

Results are added to the object.

More details on the equations used can be found in the paper Reichstein et al. (2005).

```{r Gebesee Reichstein Partitioning, message=FALSE, echo=TRUE}
EProcDEGeb$sMRFluxPartitionUStarScens()
```

---

## Step 5-1-2: Plotting the GPP

View the result columns. Columns [46] to [104] are added.

```{r Gebesee Extract Reichstein Results, echo=TRUE}
names(EProcDEGeb$sExportResults())
```

# Plot the GPP and Reco

Plot the GPP and Reco for `U50` scenario against time for two days (`48*2`).

```{r Gebesee Reichstein GPP and Reco Time Series, echo=TRUE}
nRec = 48*2 
plot(head(DEGebExampleV1$DateTime, nRec), 
     head(EProcDEGeb$sExportResults()$GPP_U50_f, nRec), type = "l",
     xlab = 'Time', ylab = 'NEE_U50')
lines(head(DEGebExampleV1$DateTime, nRec), 
      head(EProcDEGeb$sExportResults()$Reco_U50,nRec), type = "l", lty = 2, col = 'red')
```

---

## Step 5-1-3: Gebesee Lasslop Partitioning

Partitioning the data into the  fractions of the Gross Primary Production (GPP) and ecosystem respiration ($R_{eco}$) using all $u_*$ scenarios. This uses the 'sGLFluxPartitionUStarScens` function. 

Results are added to the object.

More details on the equations used can be found in the Lasslop et al. (2010).

---

```{r Gebesee Lasslop Partitioning, message=FALSE, echo=TRUE}
EProcDEGeb$sGLFluxPartitionUStarScens()
```

---

View the result columns. Columns [105] to [140] are added.

```{r Gebesee Extract Lasslop Results, echo=TRUE}
names(EProcDEGeb$sExportResults())
```

---

Plot the GPP and Reco for `U50` scenario against time for two days (`48*2`).

```{r Gebesee Lasslop GPP and Reco Time Series, echo=TRUE}
nRec <- 48*2 
plot(head(DEGebExampleV1$DateTime, nRec), 
     head(EProcDEGeb$sExportResults()$GPP_U50_f,nRec), 
     type = "l", xlab = 'Time', ylab = 'NEE_U50')
lines(head(DEGebExampleV1$DateTime, nRec), 
      head(EProcDEGeb$sExportResults()$Reco_U50,nRec), 
      type = "l", lty = 2, col = 'red')
```

---

## Step 5-1-4: Fingerprint Plots of GPP_DT and Reco_DT 

The fingerprint plots can be plotted for the GPP and $R_{eco}$.

```{r Gebesee Fingerplots of GPP_DT and Reco_DT, echo=TRUE}
EProcDEGeb$sPlotFingerprintY('GPP_DT_U50', Year = 2006)
EProcDEGeb$sPlotFingerprintY('Reco_DT_U50', Year = 2006)
```

---

## Step 5-1-5: Export the Gebesee Results

This part will produce a text file for analysis outside of R. It will be placed in the folder `results`.

```{r Export Gebesee Results, echo=TRUE}
GebData <- EProcDEGeb$sExportData() # Write the original data to GebData.
GebResults <- EProcDEGeb$sExportResults() # Write the results of REddyProc to GebResults.
GebCombResults <- cbind(GebData, GebResults)
fWriteDataframeToFile(GebCombResults, "DE-Geb_Part.txt", Dir = "../results")
```

---

# Step 6-1: Gebesee: Bias with u*-Threshold

## Calculating the Bias for the Year 2004

We will be calculating the bias of NEE due to the $u_*$-threshold for 2004.

Check the names of the columns of `GebCombResults`.

```{r Check Names of GebCombResults, echo=TRUE}
names(GebCombResults)
```


# Create a Factor Column to Distinguish the Year

First, create an integer column `year`.

```{r Create year, echo=TRUE}
GebCombResults$year <- as.POSIXlt(GebCombResults$DateTime)$year + 1900
str(GebCombResults$year)
```

---

Create a subset data frame from the combined results. 

```{r Gebesee Bias 2004, echo=TRUE}
Geb2004 <- subset(GebCombResults, year == 2004)
```

# Visualize the Difference of NEE Among the Scenarios

Using a box plot, we can see the changes.

```{r Gebesee Boxplots, echo=TRUE}
boxplot(Geb2004$NEE_U05_f, Geb2004$NEE_U50_f, Geb2004$NEE_U95_f, outline = FALSE)
boxplot(Geb2004$NEE_U05_f, Geb2004$NEE_U50_f, Geb2004$NEE_U95_f, outline = FALSE, 
        ylim = c(0,1.5))
```

# Step 6-1-1: Calculate the Annual Mean of NEE for each u*-Threshold Scenario for 2004

We will use the gap-filled 2004 data of the difference scenarios.

Create a variable that contains the means of the different scenarios: `U05`, `U50`, and `U95`.

```{r Gebesee Mean, echo=TRUE}
GebScenarios <- c("uStar","U05","U50","U95")
NEE_UStar <- sapply(GebScenarios, function(suffix){
  colName = paste0("NEE_",suffix,"_f")
  mean(Geb2004[[colName]])
})
NEE_UStar
```

## Step 6-1-2: Calculate the Statistics

Calculate the mean, standard deviation, and relative error.

```{r Gebesee Calculate Statistics, echo=TRUE}
c(mean(NEE_UStar), sd(NEE_UStar), sd(NEE_UStar)/abs(mean(NEE_UStar)))
```

# Step 7-1: Random Uncertainty Aggregation

## Step 7-1-1: Gebesee Calculate Error Terms

To calculate the error, the replaced NEE, the NEE calculated using the gap-filling method or `NEE_uStar_fall`, is subtracted from the original NEE values `NEE_ustar_orig`. The resulting value is the residual.

The original number of non-bootstrapped data for all and 2004.

```{r Gebesee Number of Original Data, echo=TRUE}
n_all <- sum(GebCombResults$NEE_uStar_fqc == 0) 
n_all
n_2004 <- sum(Geb2004$NEE_uStar_fqc == 0)
n_2004
```

The residuals are calculated for all the results and the year 2004 for comparison.

```{r Gebesee Calculate Residuals, echo=TRUE}
GebCombResults$residual <- ifelse(GebCombResults$NEE_uStar_fqc == 0,
                                  GebCombResults$NEE_uStar_orig - GebCombResults$NEE_uStar_fall,
                                  NA)

Geb2004$residual <- ifelse(Geb2004$NEE_uStar_fqc == 0,
                           Geb2004$NEE_uStar_orig - Geb2004$NEE_uStar_fall,
                           NA)
```

# Step 6-1-2: Calculate the Empirical Autocorrelation Function

Calculate the effective autocorrelation components.

```{r Gebesee Autocorrelation, echo=TRUE}
library(lognorm)
rho_all <- computeEffectiveAutoCorr(GebCombResults$residual)
plot(rho_all[-1], ylab = 'rho all', pch = 19)

rho_2004 <- computeEffectiveAutoCorr(Geb2004$residual)
plot(rho_2004[-1], ylab = 'rho 2004', pch = 19)
```

# Step 6-1-3: Calculate the Effective Number of Observations

We can calculate the number by using the autocorrelation function.
Create the variable `nEff_all` and compare to the number of good observations `n_all`.

```{r Gebesee Calculate the Effective Number of Observation, echo=TRUE}

nEff_all <- computeEffectiveNumObs(GebCombResults$residual, na.rm = TRUE, 
                                   effAcf = rho_all)
c(nEff_all, n_all)
```

# Step 6-1-4: Calculate the Effective Number of Observation for 2004 

For 2004, create the variable `nEff_2004` and compare to the number of good observations `n_2004`.

```{r Gebesee Effective Number for 2004, echo=TRUE}
nEff_2004 <- computeEffectiveNumObs(Geb2004$residual, na.rm = TRUE, 
                                    effAcf = rho_2004) 
c(nEff_2004, n_2004)
```

# Step 6-1-5: Calculate the Mean Annual NEE and Standard Deviation for 2004

Using the non-gap-filled data (`NEE_Ustar_f`), the relative error can be calculated. 

Do not use gap-filled records in the uncertainty estimation here.

The mean, standard deviations, and covariance.

```{r Gebesee NEE and SD for 2004, echo=TRUE}

NEE_notGapFilled <- mean(Geb2004$NEE_uStar_f)

sd_notGapFilled <- Geb2004$NEE_uStar_fsd[Geb2004$NEE_uStar_fqc == 0]

sdNEE_notGapFilled = sqrt(mean(sd_notGapFilled^2)) / sqrt(nEff_all - 1)

c(mean = NEE_notGapFilled, sd = sdNEE_notGapFilled, cv = sdNEE_notGapFilled/abs(NEE_notGapFilled))
```

## Step 6-1-6: Combined Uncertainties for the $u_*$-Thresholds and Random Uncertainties

Calculate the combined uncertainties of the: 

1. NEE for different $u_*$ scenarios.
2. NEE not gap-filled.

The combined uncertainties.

```{r Gebesee Combined Uncertainty, echo=TRUE}
sdNEEUStar <- sd(NEE_UStar)
sdNEECombined <- sqrt(sdNEEUStar^2 + sdNEE_notGapFilled^2)
sdNEECombined 
```

# References

Lasslop G, Reichstein M, Papale D, et al. (2010) Separation of net ecosystem exchange into assimilation and respiration using a light response curve approach: critical issues and global evaluation. Global Change Biology, Volume 16, Issue 1, Pages 187-208

Reichstein M, Falge E, Baldocchi D et al. (2005) On the separation of net ecosystem exchange into assimilation and ecosystem respiration: review and improved algorithm. Global Change Biology, 11, 1424-1439.
